---
title: "Día_2"
output: 
  html_document:
    toc: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)
```

```{r}
library(ggplot2)
library(dplyr)
```

# EJEMPLOS GRÁFICAS #####

## scatter plot: ######
```{r}
Fig1 <- ggplot(data = mpg) + geom_point(aes(x = displ, y = hwy))
Fig1
Fig1+theme_bw()
```

## Example scatter plot with regression line
```{r}
Fig2 <- Fig1 +  geom_smooth(aes(x = displ, y = hwy),
                            method = 'lm', formula = y ~ x)
c_h <- round(cor(mpg$displ, mpg$hwy, method = 'pearson'),2)
Fig2 <- Fig2 + geom_label(aes(x = 3, y = 45),
                          label = paste0('Corr:',as.character(c_h)), 
                          show.legend = F, fill = 'white', color = 'blue')
Fig2
```

## Example scatter plot with color
```{r}
Fig3 <- ggplot(data = mpg) + geom_point(aes(x = displ,
                                            y = hwy, color= class))

Fig3
```

## Example scatter plot by facets
```{r}
Fig4 <- Fig3 + facet_wrap(~ class, nrow = 3)
Fig4
```

## Example scatter plot by facets – remove legend
```{r}
Fig5 <- Fig4 + theme(legend.position = 'none')
Fig5
```

## Example scatter plot by facets – change the name of the axes and title
```{r}
Fig6 <- Fig5+
  xlab('Engine displacement') + ylab('Highway miles per gallon') + 
  ggtitle ('Comsuption vs Displacement')

Fig6
```

## Example scatter plot by facets – change the margin of the text
```{r}
Fig7 <- Fig6 + theme(plot.title = element_text(hjust = 0.5,
                                               margin=margin(0,0,10,0),
                                               size = 11),
                     axis.text.x = element_text(margin=margin(0,0,10,0)),
                     axis.text.y = element_text(margin=margin(0,0,0,10)))

Fig7
```

## Example scatter plot by facets – change style of the grid
```{r}
Fig8 <- Fig7 + theme(panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank())

Fig8
```

## Example scatter plot by facets – change the colors manually and save picture
```{r}
Fig9 <- Fig8 + scale_color_manual(values = c('red', 'blue', 'orange',
                                             'black', 'yellow',
                                             'pink', 'green'))
Fig9

#para guardar la figura
# ggsave(filename = paste0(file.path('output','test.png')), height = 10, 
#        width = 10 , units = "cm", dpi = 500)
```

##boxplot ########

### Box-plot of engine-displacement by class of cars
```{r}
Fig10 <- ggplot(data = mpg, aes(class,displ)) + 
  geom_boxplot(color="red", fill="orange",
               alpha=0.2, width = 0.8) +
  stat_boxplot(geom = 'errorbar', color="red", width = 0.4)+
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5,
                                  margin=margin(0,0,10,0),
                                  face = 'bold'),
        axis.ticks.x = element_blank()) +
  xlab('Class') + ylab('Engine displacement') +
  ggtitle('Boxplot- Engine displacement')

Fig10
```

### Box-plot of engine-displacement by class of cars
```{r}
library(viridis)
Fig11 <- ggplot(data = mpg, aes(x=class,y=displ,fill=class)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5, margin=margin(0,0,10,0),
                                  face = 'bold'), axis.ticks.x = element_blank(),   legend.position="none") + 
  xlab('Class') + ylab('Engine displacement') + ggtitle('Boxplot- Engine displacement')
Fig11
```

## violin plot ######
```{r}
Fig12 <- ggplot(data = mpg, aes(class,displ)) + 
  geom_violin(color="red", fill="orange", alpha=0.2)+ theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, margin=margin(0,0,10,0), face = 'bold'),
        axis.ticks.x = element_blank()) +
  xlab('Class') + ylab('Engine displacement') + ggtitle('Violinplot- Engine displacement')

Fig12
```

## barplot #######
```{r}
Fig14 <- ggplot(data = mpg) + geom_bar(mapping = aes(x = class), width = 0.5, 
                                       color = 'blue', fill='royalblue') +
  theme(plot.title = element_text(hjust = 0.5,margin=margin(0,0,10,0), size = 11),
        axis.text.x = element_text(margin=margin(0,0,10,0)), axis.text.y = element_text(margin=margin(0,0,0,10)), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(colour = "black", size = 1)) +
  xlab('Class') + ylab('Count') + ggtitle('Frequency of each class')

Fig14
```

```{r}
Fig15 <- ggplot(data = mpg) + geom_bar(mapping = aes(x = class, fill = trans), 
                                       position = 'dodge' ) +
  theme(plot.title = element_text(hjust = 0.5,margin=margin(0,0,10,0), size = 11),
        axis.text.x = element_text(margin=margin(0,0,10,0)), axis.text.y = element_text(margin=margin(0,0,0,10)), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(colour = "black", size = 1)) +
  xlab('Class') + ylab('Count') + ggtitle('Frequency of each class and type of transmission')
Fig15
```

```{r}
Fig16 <- ggplot(data = mpg) + 
  geom_bar(mapping = aes(x = class, fill = trans), position = 'stack' ) +
  theme(plot.title = element_text(hjust = 0.5,margin=margin(0,0,10,0), size = 11),
        axis.text.x = element_text(margin=margin(0,0,10,0)), axis.text.y = element_text(margin=margin(0,0,0,10)), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(colour = "black", size = 1)) +
  xlab('Class') + ylab('Count') + ggtitle('Frequency of each class and type of transmission')
Fig16
```

```{r}
Fig17 <- ggplot(data = mpg) + geom_bar(mapping = aes(x = class, fill = trans), 
                                       position = 'stack' ) + coord_flip() +
  theme(plot.title = element_text(hjust = 0.5,margin=margin(0,0,10,0), size = 11),
        axis.text.x = element_text(margin=margin(0,0,10,0)), axis.text.y = element_text(margin=margin(0,0,0,10)), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(colour = "black", size = 1)) +
  xlab('Class') + ylab('Count') + ggtitle('Frequency of each class and type of transmission')
Fig17
```

```{r}
mpg$year <- factor(mpg$year, levels= c('1999', '2008'))
Fig18 <- ggplot(data = mpg)+ geom_tile(aes(x = year, y = manufacturer, fill = hwy),colour='black') +  theme_minimal() +
  theme(axis.title.x=element_blank(),axis.title.y=element_blank(),
        axis.text.x = element_text(hjust = 0.9, vjust = 0.9, angle = 45)) +
  scale_fill_gradient(low="red", high="blue")
Fig18
```



# Ejercicio 1 #####

```{r}
datos<- read.csv("C:/Users/enriq/Desktop/Master/Data Science (MIB)/Clases Juan/Día 2/data/season-1819_csv.csv", header = T, sep = ",")

construir_tabla <- function(datos,equipo) {
  win_home <- datos %>%
    filter(HomeTeam == equipo & FTR == "H")
  win_home <- nrow(win_home)
  no_win_home <- datos %>%
    filter(HomeTeam == equipo & FTR != "H")
  no_win_home <- nrow(no_win_home)
  win_away <- datos %>%
    filter(AwayTeam == equipo & FTR == "A")
  win_away <- nrow(win_away)
  no_win_away <- datos %>%
    filter(AwayTeam == equipo & FTR != "A")
  no_win_away <- nrow(no_win_away)
  
  tabla <- matrix(nrow = 2, ncol = 2)
  tabla[1,1] <- win_home
  tabla[1,2] <- win_away
  tabla[2,1] <- no_win_home
  tabla[2,2] <- no_win_away
  
  colnames(tabla) <- c("Home","Away")
  rownames(tabla) <- c("Win", "No Win")
  
  return(tabla)
}

tabla_real <- construir_tabla(datos,"Sociedad")
tabla_real
fisher.test(tabla_real)

tabla_barsa <- construir_tabla(datos,"Barcelona")
tabla_barsa
fisher.test(tabla_barsa)

cat("En resumen, ni el Barça ni la Real Sociedad dependen de estar en casa para ganar")
```

# Ejercicio 2 #####

```{r}
data <- read.csv("C:/Users/enriq/Desktop/Master/Data Science (MIB)/Clases Juan/Día 2/data/crimeUS.csv", header = T, sep = ",")

data1 <- data %>%
  aggregate(cbind(afam, cauc) ~ state, FUN = mean)

data1$afam <- round(data1$afam)
data1$cauc <- round(data1$cauc)


data2 <- data %>%
  group_by(state) %>%
  summarise(African_American = round(mean(afam)),
            Caucasic = round(mean(cauc)))

head(data2)
head(data1)




crear_tabla <- function(estado1, estado2, datos) {
  
  V1 <- datos %>%
    filter(state == estado1) %>%
    pull(cauc) %>%
    sum()
  V2 <- datos %>%
    filter(state == estado1) %>%
  pull(afam) %>%
    sum()
  V3 <- datos %>%
    filter(state == estado2) %>%
  pull(cauc) %>%
    sum()
  V4 <- datos %>%
    filter(state == estado2) %>%
  pull(afam) %>%
    sum()
  
  tabla <- matrix(nrow = 2, ncol = 2)
  tabla[1,1] <- V1
  tabla[1,2] <- V2
  tabla[2,1] <- V3
  tabla[2,2] <- V4
  
  colnames(tabla) <- c("cauc","afam")
  rownames(tabla) <- c(estado1, estado2)
  
  return(tabla)
}

tabla1 <- crear_tabla("District of Columbia","Hawaii",data)
tabla1
fisher.test(tabla1)

tabla2 <- crear_tabla("Florida","Hawaii",data)
tabla2
fisher.test(tabla2)
```


# Ejercicio 3 #####

```{r}
data <- read.csv("C:/Users/enriq/Desktop/Master/Data Science (MIB)/Clases Juan/Día 2/data/crimeUS.csv", header = T, sep = ",")

data1 <- data %>%
  mutate(violent_crimes = (violent/population)*100000)

comparar <- function(s1, s2, s3, data) {
data2 <- data %>%
  group_by(state) %>%
  filter(state %in% c(s1, s2, s3))

modelo <- aov(data2$violent_crimes ~ data2$state)
summary(modelo)

bp <- ggplot(data2, aes(x = state, y = violent_crimes)) + 
  geom_boxplot(color = "blue", fill = "blue", alpha = 0.3) +
  theme_minimal()
bp

#return(summary(modelo))
return(bp)
}

comp1 <- comparar("Alaska","Georgia","Missouri", data1)
comp1

comp2 <- comparar("District of Columbia","Hawaii","Idaho", data1)
comp2
```